Delivering Reliable Software in the European Scientific Arena: Evolution of Methodologies and New Trend Analysis

### 2018-02-04 paper-ieee-trans.tex
Add brief explaination of SQA requirements

Three main pillars sustain the project's strategy to seek reliability in the 
software produced. A solid \textit{Software Quality Assurance (SQA) criteria
and metrics gathering definition} guides the development of the software in
a per-change basis, validating the software at the very early stages. The 
software delivery process is designed as a bottom-up \textit{adoption of DevOps
principles}, passing through Continuous Integration (CI) and Continuous
Delivery (CD) approaches. Once the software is released, it follows a 
\textit{two-factor validation} in the project's pilot infrastructures and
resource provider's early-adoption.

@@

The software quality policies \cite{indigo-d31} cover: 1) the identification
and description of the \emph{quality requirements} that the software need to
comply with, and 2) the \emph{quality metrics}.

The requirements are checked at early stages in the software development
process, integrated in the Jenkins CI service \cite{indigo-jenkins}, so that any
bug or design issue is likely to be detected and corrected promptly. The source
code is publicly available in GitHub repositories under the project organization called
{\sl indigo-dc} \cite{indigo-github} to increase the visibility of the products
catalogue, promoting the external contributions and software adoption. Furthermore,
the source code is checked for compliance with community de--facto style standards,
selected for each one of the variety of programming languages being used as seen in
Figure \ref{fig:fig_codestyle}.

@@

\subsubsection{Continuous Integration}
The INDIGO--DataCloud project promoted the application of a CI scenario to enforce
the SQA requirements -- described in Section \ref{subsec:sqa} -- for any piece of
software being produced within the project. Such environment requires an automation
ready--to--go infrastructure where the different technologies involved -- source code
management platform, automation server and containerization -- interact with each
other to trigger the source code validation pipeline for each candidate change. The
pipeline is comprised of diverse quality checks, enforcing at least the successful
execution of code style and minimum unit testing coverage. To take over the 
implementation of such infrastructure, the project leveraged on tightly integrated open
source tools such as GitHub \cite{github}, as the online source code repository,
and Jenkins \cite{jenkins}, the event-response CI system. Docker container provisioning
is used to provide instant computing power needed for executing the pipeline jobs.

>>>>

In the following section we will describe the three main pillars that sustain the project's strategy to seek reliability in the 
software produced. A solid \textit{Software Quality Assurance (SQA) criteria
and metrics gathering definition} guides the development of the software in
a per-change basis, validating the software at the very early stages. The 
software delivery process is designed as a bottom-up \textit{adoption of DevOps
principles}, passing through Continuous Integration (CI) and Continuous
Delivery (CD) approaches. Once the software is released, it follows a 
\textit{two-factor validation} in the project's pilot infrastructures and
resource provider's early-adoption.

@@

The software quality policies \cite{indigo-d31} cover: 1) the identification
and description of the \emph{SQA requirements} that the produced software need 
to comply with, and 2) the \emph{quality metrics}, to monitor each software product's 
behaviour throughout the development, release and production stages.

The SQA requirements are checked automatically at early stages in the software
development process, following a CI approach described in Section \ref{subsec:ci}, so
that that any bug or design issue is likely to be detected and corrected promptly.

\subsubsection{Code style}
The source code is adhered to any style standard and shall be compliant with the de-facto
guidelines present for the programming language being used.
\subsubsection{Unit and Functional testing}
Whenever a new change involves a new functionality addition, it is requested to be tested. 
Regression testing in this context is done enforcing the successful execution of the previous
functionality tests already present in the code. On the other hand, unit testing evaluates the
code's internal design setting a threshold of 70\% coverage for the code developed.
\subsubsection{Documentation}
\subsubsection{Code review}
Represents the last step in the change management pipeline, once the candidate change has 
successfully passed through the testing methods described previously. It implies the human-based
revision of the proposed change to discuss its adequance in terms of scope, objective fulfillment, 
documentation completeness, etc. On approval, the candidate change is definetively merged into
production.
\subsubsection{Automated deployment}

integrated in the Jenkins CI service \cite{indigo-jenkins}, so that any
bug or design issue is likely to be detected and corrected promptly. The source
code is publicly available in GitHub repositories under the project organization called
{\sl indigo-dc} \cite{indigo-github} to increase the visibility of the products
catalogue, promoting the external contributions and software adoption. Furthermore,
the source code is checked for compliance with community de--facto style standards,
selected for each one of the variety of programming languages being used as seen in
Figure \ref{fig:fig_codestyle}.

@@

\subsubsection{Continuous Integration}
\label{subsec:ci}
The INDIGO--DataCloud project promoted the application of a CI scenario to enforce
the SQA requirements -- described in Section \ref{subsec:sqa} -- for any piece of
software being produced within the project. Such environment requires an automation
ready--to--go infrastructure where the different technologies involved -- source code
management platform, automation server and containerization -- interact with each
other to trigger the source code validation pipeline for each candidate change. The
pipeline is comprised of diverse quality checks, enforcing at least the successful
execution of code style and minimum unit testing coverage. To take over the 
implementation of such infrastructure, the project leveraged on tightly integrated open
source tools such as GitHub \cite{github}, as the online source code repository,
and Jenkins \cite{jenkins}, the event-response CI system. Docker container provisioning
is used to provide instant computing power needed for executing the pipeline jobs.
